TBD

### 2019-02-06 inverse_problems/inverse_problems.tex
updates on the Poisson process

\newcommand{\LL}{\mathcal{L}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\ep}{\varepsilon}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\GCD}{\mathbf{GCD}}
\newcommand{\XX}{\mathcal{X}}
\newcommand{\SUM}{\text{sum}}
\newcommand{\1}{\mathbf{1}}
\newcommand{\rr}{\textbf{r}}
\newcommand{\ii}{\textbf{i}}
\newcommand{\jj}{\textbf{j}}

\newcommand{\II}{\mathcal{I}}
\newcommand{\kk}{\textbf{k}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\mb}{\mathbf}
\newcommand{\mk}{\mathfrak}
\newcommand{\mc}{\mathcal}
\newcommand*\Bell{\ensuremath{\boldsymbol\ell}}
\newcommand{\TODO}[1]{{\color{red}{[#1]}}}
\newcommand{\revised}[1]{{\color{blue}{#1}}}
\newcommand{\R}{\mathbb{R}}

@@

In this paper, we consider the problem of recovering signals that appear 
multiple times at unknown locations in a noisy measurement. 
In particular, we consider three/four models.

\paragraph{The homogeneous well-separated model.}
%This problem can be conveniently posed as  \emph{blind deconvolution}.
Let $x\in\RL$ be the sought signal and let $y\in\RN$ be the observed data, where we assume $N$ is  far larger than $L$. Let  $s \in \{0, 1\}^{N-L+1}$ be a binary signal indicating (with 1's) the starting positions of all occurrences of $x$ in $y$, so that, with additive white Gaussian noise:
\begin{equation} 
y  =  x \ast s + \varepsilon, \quad  \varepsilon   \sim \mathcal{N}(0,\sigma^2 I_N),
\label{eq:model}
\end{equation}
where $\ast$ denotes linear convolution. 
While both $x$ and $s$ are unknown, the goal is only to estimate $x$ from $y$. 
This is a key difference with other works on  \emph{blind deconvolution}, a longstanding problem arising in a variety of engineering and scientific applications such as astronomy, communication, image deblurring, system identification and optics; see~\cite{jefferies1993restoration,shalvi1990new,ayers1988iterative,abed1997blind}, just to name a few. 
Different variants of the blind deconvolution problem have been subject recently to a through  analysis, focusing on low noise regimes~\cite{ahmed2014blind,li2016identifiability,li2016rapid,lee2017blind,ling2017blind,kuo2019geometry}.
The parameters of the signal $s$ (the locations of its nonzero values) are the \emph{nuisance variables} of the problem. 
The model~\eqref{eq:model} appears in several scientific applications, including spike sorting~\cite{lewicki1998review}, passive radar~\cite{gogineni2017passive} and system identification~\cite{ljung1998system}.

@@

\paragraph{The homogeneous Poisson model.}
In this model, we assume that the support signal $s$ is generated by a Poisson process. Bla bla bla. While this model is very different from the well-separated model, in Section~\ref{sec:support} we show that TKTK

\paragraph{The heterogeneous well-separated model.}

The homogeneous well-separated model can be naturally generalized to the problem of estimating a set of $K$ signals $x_1,\ldots,x_K\in\RL$ from their mixed blind deconvolution
\begin{equation} 
y  =  \sum_{i=1}^Kx_i \ast s_i + \varepsilon, \quad  \varepsilon   \sim \mathcal{N}(0,\sigma^2 I_N),
\end{equation}
where again $s_i$ denotes a binary signal indicating the support of the occurrences of $x_i$ in the data. We say that the separation condition is met \TODO{to verify}
\begin{eqnarray}
\text{If } s:=\bigcup\limits_{i=1}^Ks_i \text{ satisfies}~\eqref{eq:spacing} \text{ and for any } i\neq j, \thinspace s_i\cap s_j=\emptyset.
\end{eqnarray}

@@

\paragraph{The heterogeneous Poisson model?.}

@@

\section{Autocorrelation analysis} \label{sec:AC_analysis}

For a signal $z\in \R^m$, the autocorrelation of order $q = 1, 2, \ldots$ is given for any integer shifts $\ell_1, \ldots, \ell_{q-1}$ by
\begin{equation}
a_z^q[\ell_1,\ldots,\ell_{q-1}]   = \frac{1}{m} \sum_{i=-\infty}^{\infty} z[i]z[i+\ell_1]\cdots z[i+\ell_{q-1}],
\label{eq:ac_general}
\end{equation}
where indexing of $z$ out of the range $0, \ldots, m-1$ is zero-padded.
Explicitly, the first-, second- and third-order autocorrelations are given by: 
\begin{align} 
a_z^1 & = \frac{1}{m} \sum_{i=0}^{m-1} z[i], \nonumber\\
a_z^2[\ell] & = \frac{1}{m} \sum_{i = \max\{0, -\ell\}}^{m-1 + \min\{0, -\ell\}} z[i]z[i+\ell],\\
a_z^3[\ell_1,\ell_2] & = \frac{1}{m} \sum_{i = \max\{0, -\ell_1, -\ell_2\}}^{m-1 + \min\{0, -\ell_1, -\ell_2\}} z[i]z[i+\ell_1]z[i+\ell_2].  \nonumber \label{eq:ac_special}
\end{align}
The autocorrelation functions have symmetries. Specifically, $$a_z^2[\ell] = a_z^2[-\ell],$$ and
$$a_z^3[\ell_1,\ell_2] = a_z^3[\ell_2,\ell_1]=a_z^3[-\ell_1,\ell_2-\ell_1].
$$
For our purposes, this will be applied both to $x$ (of length $L$) and to $y$ (of length $N$).

\TODO{Start directly with heterogeneous model, but first need to verify that the Poisson equivalence is true holds also for this case.}
Under the separation condition, the relation between autocorrelations of the micrograph and those of $x$ is particularly simple, as we now show. It is useful to introduce some notation: let $M$ denote the number of occurrences of $x$ in $y$ (that is, the number of 1's in $s$), and let
\begin{equation}
\gamma  = \frac{M L}{N}
\label{eq:gamma}
\end{equation}
denote the density of $x$ in $y$ (that is, the fraction of entries of $y$ occupied by occurrences of $x$.) The separation condition %~\eqref{eq:spacing} 
imposes $\gamma\leq\frac{L}{2L-1}\approx 1/2$.

@@

@@

>>>>

\newcommand{\LL}{\mathcal{L}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\ep}{\varepsilon}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\GCD}{\mathbf{GCD}}
\newcommand{\XX}{\mathcal{X}}
\newcommand{\SUM}{\text{sum}}
\newcommand{\1}{\mathbf{1}}
\newcommand{\rr}{\textbf{r}}
\newcommand{\ii}{\textbf{i}}
\newcommand{\jj}{\textbf{j}}
\newcommand{\Poisson}{\text{Poisson}}
\newcommand{\II}{\mathcal{I}}
\newcommand{\kk}{\textbf{k}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\mb}{\mathbf}
\newcommand{\mk}{\mathfrak}
\newcommand{\mc}{\mathcal}
\newcommand*\Bell{\ensuremath{\boldsymbol\ell}}
\newcommand{\TODO}[1]{{\color{red}{[#1]}}}
\newcommand{\revised}[1]{{\color{blue}{#1}}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\M}{\mathcal{M}}
\renewcommand{\P}{\mathbb{P}}
\renewcommand{\L}{\mathcal{L}}

@@

In this paper, we consider the problem of recovering signals that appear 
multiple times at unknown locations in a noisy measurement.
Let $x\in\RL$ be the sought signal and let $y\in\RN$ be the observed data, where we assume $N$ is  far larger than $L$. Let  $s \in \{0, 1\}^{N-L+1}$ be a binary signal indicating (with 1's) the starting positions of all occurrences of $x$ in $y$, so that, with additive white Gaussian noise:
\begin{equation} 
y  =  x \ast s + \varepsilon, \quad  \varepsilon   \sim \mathcal{N}(0,\sigma^2 I_N),
\label{eq:model}
\end{equation}
where $\ast$ denotes linear convolution. 
While both $x$ and $s$ are unknown, the goal is only to estimate $x$ from $y$. 
This is a key difference with other works on  \emph{blind deconvolution}, a longstanding problem arising in a variety of engineering and scientific applications such as astronomy, communication, image deblurring, system identification and optics; see~\cite{jefferies1993restoration,shalvi1990new,ayers1988iterative,abed1997blind}, just to name a few. 
Different variants of the blind deconvolution problem have been subject recently to a through  analysis, focusing on low noise regimes~\cite{ahmed2014blind,li2016identifiability,li2016rapid,lee2017blind,ling2017blind,kuo2019geometry}.
The parameters of the signal $s$ (the locations of its nonzero values) are the \emph{nuisance variables} of the problem. 
The model~\eqref{eq:model} appears in several scientific applications, including spike sorting~\cite{lewicki1998review}, passive radar~\cite{gogineni2017passive} and system identification~\cite{ljung1998system}.

@@

\subsection{Extensions}

\paragraph{The heterogeneous well-separated model.}
The homogeneous well-separated model can be naturally generalized to the problem of estimating a set of $K$ signals $x_1,\ldots,x_K\in\RL$ from their mixed blind deconvolution
\begin{equation} 
y  =  \sum_{i=1}^Kx_i \ast s_i + \varepsilon, \quad  \varepsilon   \sim \mathcal{N}(0,\sigma^2 I_N),
\end{equation}
where again $s_i$ denotes a binary signal indicating the support of the occurrences of $x_i$ in the data. We say that the separation condition is met \TODO{to verify}
\begin{eqnarray}
\text{If } s:=\bigcup\limits_{i=1}^Ks_i \text{ satisfies}~\eqref{eq:spacing} \text{ and for any } i\neq j, \thinspace s_i\cap s_j=\emptyset.
\end{eqnarray}

@@

\paragraph{Poisson model.}

Until now we have presented models assuming that the signals' occurrences are well separated. This condition can be alleviated by assuming a Poisson distribution on $s$.

Let $X\in\RL$ be a random vector drawn from some fixed distribution. 
We observe a random vector $Y\in\R^N$ generated as follows. Points are chosen in $\{0,\dots,N-L\}$ according to a Poisson process with parameter $\gamma N$. 
The parameter $\gamma$ plays the role of the density of the signal in the measurement. 
For each point $i$ that is chosen from $1$ to $N-L$, a random vector $X$ from the distribution is then placed in the large vector, with element $0$ at location $i$, with overlapping vectors being added together. 
If the distribution of $X$ is a single Dirac function, then we observe multiple repetition of the same signals, like in the homogeneous case. If the distribution consists of $K$ Diracs, then it is a heterogeneous model.

If $M_i$ denotes the number of hits at location $i$, $1 \le i \le N$, then by definition of the Poisson process $M_i$'s are i.i.d.\ and $M_i \sim \Poisson(\gamma)$. Conditional on the value of $M = (M_1,\dots,M_N)$, if we let $X_1^{i},\dots,X_{M_i}^i$ denote the random vectors with position 0 located at $i$, then $X_{k_1}^{i}$ and $X_{k_2}^{i}$ are independent for $k_1 \ne k_2$.
With this notation,  we can write each entry as:
%
\begin{align}
%
Y[i] = \sum_{j=0}^{L-1} \sum_{k=1}^{M_{i-j}} X_k^{i-j}[j].
%
\end{align}

@@

\section{Autocorrelation analysis} \label{sec:AC_analysis}
\TODO{Maybe we should re-write this section for a random signal and then particularized to the homogeneous and heterogeneous cases}
For a signal $z\in \R^m$, the autocorrelation of order $q = 1, 2, \ldots$ is given for any integer shifts $\ell_1, \ldots, \ell_{q-1}$ by
\begin{equation}
a_z^q[\ell_1,\ldots,\ell_{q-1}]   = \frac{1}{m} \sum_{i=-\infty}^{\infty} z[i]z[i+\ell_1]\cdots z[i+\ell_{q-1}],
\label{eq:ac_general}
\end{equation}
where indexing of $z$ out of the range $0, \ldots, m-1$ is zero-padded.
Explicitly, the first-, second- and third-order autocorrelations are given by: 
\begin{align} 
a_z^1 & = \frac{1}{m} \sum_{i=0}^{m-1} z[i], \nonumber\\
a_z^2[\ell] & = \frac{1}{m} \sum_{i = \max\{0, -\ell\}}^{m-1 + \min\{0, -\ell\}} z[i]z[i+\ell],\\
a_z^3[\ell_1,\ell_2] & = \frac{1}{m} \sum_{i = \max\{0, -\ell_1, -\ell_2\}}^{m-1 + \min\{0, -\ell_1, -\ell_2\}} z[i]z[i+\ell_1]z[i+\ell_2].  \nonumber \label{eq:ac_special}
\end{align}
The autocorrelation functions have symmetries. Specifically, $$a_z^2[\ell] = a_z^2[-\ell],$$ and
$$a_z^3[\ell_1,\ell_2] = a_z^3[\ell_2,\ell_1]=a_z^3[-\ell_1,\ell_2-\ell_1].
$$
For our purposes, this will be applied both to $x$ (of length $L$) and to $y$ (of length $N$).

Under the separation condition, the relation between autocorrelations of the micrograph and those of $x$ is particularly simple, as we now show. It is useful to introduce some notation: let $M$ denote the number of occurrences of $x$ in $y$ (that is, the number of 1's in $s$), and let
\begin{equation}
\gamma  = \frac{M L}{N}
\label{eq:gamma}
\end{equation}
denote the density of $x$ in $y$ (that is, the fraction of entries of $y$ occupied by occurrences of $x$.) The separation condition %~\eqref{eq:spacing} 
imposes $\gamma\leq\frac{L}{2L-1}\approx 1/2$.

@@

In this section we omit the effect of the noise as it remains identical to the well-separated model.

The moments of $Y$ are defined by
%
\begin{align}
%
m^1_Y[i] = \E Y[i], \quad 0 \le i \le L-1,
%
\end{align}
%
\begin{align}
%
m^2_Y[i,j] = \E Y[i] Y[j], \quad 0 \le i,j \le L-1,
%
\end{align}
%
and
%
\begin{align}
%
m^3_Y[i,j,k] = \E Y[i] Y[j] Y[k], \quad 0 \le i,j,k \le L-1.
%
\end{align}

%We will also denote by $\L_l$ the autocorrelations of $X$, $l = 1,2,3$:
%%
%\begin{align}
%%
%\L_1 = \sum_{i=0}^{L-1} \M_1[i] \in \R,
%%
%\end{align}
%%
%\begin{align}
%%
%\L_2(\Delta) = \sum_{i=0}^{L-1} \M_2[i,i+\Delta]  \in \R^L,
%%
%\end{align}
%%
%and
%%
%\begin{align}
%%
%\L_3(\Delta_1,\Delta_2) = \sum_{i=0}^{L-1} \M_3[i,i+\Delta_1,i+\Delta_2]  \in \R^{L \times L}.
%%
%\end{align}
%
%Note that in the strongly-separated model, the first three observed moments are, respectively, $\L_1$, $\L_2(\Delta)$, and $\L_3(\Delta_1,\Delta_2)$. \TODO{I think we can remove it}
%%
%
We now prove that the first three autocorrelations of the well-separated model can be recovered from the moments of the Poisson process.

\begin{proposition} \label{prop:poisson}
Let $a_X^1, a_X^2$ and $a_X^3$ be the first three autocorrelation function of the well-separated model \TODO{This is a bit unclear now, but if I get it right it holds true for any distribution over $X$}. Then, for any   $0\leq i\leq L-1$ we have 
\begin{align}
m_Y^1[i] &= \gamma a_X^1,  \nonumber \\
m_Y^2[i,i+\ell] &= (\gamma a_X^1)^2 + \gamma a_X^2[\ell], \quad \ell\in[0,L-1], \\
m_Y^3[i,i+\ell_1,i+\ell_2] &= 
(\gamma a_X^1)^3 + \gamma a_X^1  \cdot ( \gamma a^2_X[\ell_1] + \gamma a_X^2[\ell_2] + \gamma a_X^2[\ell_2-\ell_1]) + \gamma a_X^3[\ell_1,\ell_2], \quad \quad \ell_1,\ell_2\in[0,L-1]. \nonumber
\end{align}
\end{proposition}

\begin{corollary}
	From the first three moments of the Poisson process model, once can recover the first three moments from the strongly-separated model.
\end{corollary}

%%%, with the Poisson rate $\gamma$ playing the role of the ``occupancy factor''. So if recovery is possible for the strongly-separated model, it is also possible for the Poisson process model.

%
%
%\begin{rem}
%	
%	The expected values of the non-zero (for second moment) and off-diagonal (for third moment) terms are the same as without noise, as is true for the strongly-separated case. The same proof of almost sure convergence from my notes for the strongly-separated case also goes through verbatim.
%	
%	
%\end{rem}
%

@@

\section{Proof of Proposition~\ref{prop:poisson}} \label{sec:proof_prop_poisson}

\subsection{The first moment of $Y$}

To compute the first moment of $Y$, we will first condition on $M = (M_1,\dots,M_n)$, and then average over $M$. We have:
%
\begin{align}
%
\E[Y[i] | M] = \sum_{j=0}^{L-1} \sum_{k=1}^{M_{i-j}} \E X_k^{i-j}[j]
= \sum_{j=0}^{L-1} \sum_{k=1}^{M_{i-j}} \M_1[j]
= \sum_{j=0}^{L-1} M_{i-j} \M_1[j].
%
\end{align}
%
Now taking expectations over $M$ we see:
%
\begin{align}
%
\E Y[i] = \gamma \sum_{j=0}^{L-1} \M_1[j] = \gamma \L_1.
%
\end{align}

\subsection{The second moment of $Y$}

Again, we will condition on $M$ first, and then take the expectation over $M$. Fix $i_1 \ne i_2$, and let $\Delta = i_2 - i_1$. Then:
%
\begin{align}
%
Y_{i_1} Y_{i_2} 
&= \sum_{j_1=0}^{L-1} \sum_{j_2=0}^{L-1} 
\sum_{k_1=1}^{M_{i_1-j_1}}\sum_{k_2=1}^{M_{i_2-j_2}}
X_{k_1}^{i_1-j_1}[j_1] X_{k_2}^{i_2 - j_2}[j_2].
%
\end{align}

We break up the double sum over $j_1$ and $j_2$ into two terms: one where $j_2 \ne j_1 + \Delta$, and one where $j_2 = j_1 + \Delta$ or equivalently $i_1-j_1 = i_2-j_2$. In the first case, all the terms are independent, and so the expectation factors. In the second case, when $k_1 \ne k_2$ we have independence, but otherwise not. This gives (all expectations are conditional on $M$):
%
\begin{align} \label{moment2-condm}
%
\E Y_{i_1} Y_{i_2}
=& \sum_{j_1=0}^{L-1} \sum_{j_2=0}^{L-1} 
\sum_{k_1=1}^{M_{i_1-j_1}}\sum_{k_2=1}^{M_{i_2-j_2}}
\E X_{k_1}^{i_1-j_1}[j_1] X_{k_2}^{i_2 - j_2}[j_2]
\nonumber \\
=& \sum_{j_1 - j_2 \ne \Delta} \sum_{k_1} 
\sum_{k_2} \E X_{k_1}^{i_1-j_1}[j_1] X_{k_2}^{i_2 - j_2}[j_2]
\nonumber \\
& + \sum_{j_1 = 0}^{L-1} \sum_{k_1 \ne k_2} 
\E X_{k_1}^{i_1-j_1}[j_1] X_{k_2}^{i_1 - j_1}[j_1+\Delta]
\nonumber \\
& + \sum_{j_1 = 0}^{L-1} \sum_{k_1=1}^{M_{i_1-j_1}} 
\E X_{k_1}^{i_1-j_1}[j_1] X_{k_1}^{i_1-j_1}[j_1 + \Delta] 
\nonumber \\
=& \sum_{j_1 - j_2 \ne \Delta} M_{i_1-j_1} M_{i_2 - j_2} \M_1[j_1] \M_1[j_2]
\nonumber \\
& + \sum_{j_1 = 0}^{L-1} M_{i_1-j_1}(M_{i_1-j_1} - 1) \M_1[j_1] \M_1[j_1 + \Delta]
\nonumber \\
& + \sum_{j_1 = 0}^{L-1} M_{i_1-j_1} \M_2[j_1,j_1 + \Delta] .
%
\end{align}
%
Now take expectations over the Poisson random variables, using this fact:
%
\begin{lem} \label{lem-choose}
	If $M \sim \Poisson(\gamma)$, then 
	\begin{align}
	%
	\E {M\choose k} = \frac{\gamma^k}{k!}.
	%
	\end{align}
\end{lem}

We get (now the expectation is over $M$ and $X$):
%
\begin{align}
%
\E Y_{i_1} Y_{i_2} 
=& \sum_{j_1 - j_2 \ne \Delta} \E M_{i_1-j_1} M_{i_2 - j_2} \M_1[j_1] \M_1[j_2]
\nonumber \\
& + \sum_{j_1 = 0}^{L-1} \E M_{i_1-j_1}(M_{i_1-j_1} - 1) \M_1[j_1] \M_1[j_1 + \Delta]
\nonumber \\
& + \sum_{j_1 = 0}^{L-1} \E M_{i_1-j_1} \M_2[j_1,j_1 + \Delta]
\nonumber \\
=& \sum_{j_1 - j_2 \ne \Delta} \gamma^2 \M_1[j_1] \M_1[j_2]
+ \sum_{j_1 = 0}^{L-1} \gamma^2 \M_1[j_1] \M_1[j_1 + \Delta]
\nonumber \\
& + \sum_{j_1 = 0}^{L-1} \gamma \M_2[j_1,j_1 + \Delta]
\nonumber \\
=&  \bigg(\gamma \sum_{j = 0}^{L-1} \M_1[j] \bigg)^2
+ \gamma \sum_{j = 0}^{L-1} \M_2[j,j + \Delta]
\nonumber \\
=&  (\gamma \L_1)^2 + \gamma \L_2(\Delta).
%
\end{align}

But the first term in the sum is just the square of the first moment of $Y$; so from the first two moments we can recover $\gamma \L_2(\Delta)$, which is just the expected power spectrum of the random vector $X$, i.e.\ the usual second moment we have been working with.

\subsection{The third moment of $Y$}

For three distinct $i_1$, $i_2$ and $i_3$, we let $\Delta_1 = i_2 - i_1$ and $\Delta_2 = i_3 - i_1$. We have:
%
\begin{align}
%
&Y_{i_1} Y_{i_2} Y_{i_3}
\nonumber \\
&= \sum_{j_1=0}^{L-1} \sum_{j_2=0}^{L-1} \sum_{j_3=0}^{L-1} 
\sum_{k_1=1}^{M_{i_1-j_1}}\sum_{k_2=1}^{M_{i_2-j_2}} \sum_{k_3=1}^{M_{i_3-j_3}}
X_{k_1}^{i_1-j_1}[j_1] X_{k_2}^{i_2 - j_2}[j_2] X_{k_3}^{i_3 - j_3}[j_3].
\end{align}
%
We will break up the outer three sums into disjoint sums with the following ranges of indices:
%
\begin{enumerate}

	\item \label{case1}
	$j_2 = j_1 + \Delta_1$ and $j_3 = j_2 + \Delta_2 - \Delta_1$.

	\item \label{case2}
	$j_2 = j_1 + \Delta_1$ and $j_3 \ne j_2 + \Delta_2 - \Delta_1$.

	\item \label{case3}
	$j_2 \ne j_1 + \Delta_1$ and $j_3 = j_1 + \Delta_2$.

	\item \label{case4}
	$j_2 \ne j_1 + \Delta_1$ and $j_3 \ne j_1 + \Delta_2$ and $j_3 = j_2 + \Delta_2 - \Delta_1$.

	\item \label{case5}
	$j_2 \ne j_1 + \Delta_1$ and $j_3 \ne j_1 + \Delta_2$ and $j_3 \ne j_2 + \Delta_2 - \Delta_1$.

\end{enumerate}

For Case \ref{case1}, we have $\ell \equiv i_1 - j_1 = i_2 - j_2 = i_3 - j_3$. We further break up the sum:
%
\begin{align}
%
&\sum_{j=0}^{L-1} \sum_{k_1=1}^{M_\ell} \sum_{k_2=1}^{M_\ell} \sum_{k_3=1}^{M_\ell} 
X_{k_1}^{\ell}[j] X_{k_2}^{\ell}[j + \Delta_1] X_{k_3}^{\ell}[j + \Delta_2]
\nonumber \\
=& \underbrace{ \sum_{j=0}^{L-1} \sum_{k_i \text{distinct}} 
	X_{k_1}^{\ell}[j] X_{k_2}^{\ell}[j + \Delta_1] X_{k_3}^{\ell}[j + \Delta_2]
}_{\text{(a)}}
\nonumber \\
&+\underbrace{ \sum_{j=0}^{L-1} \sum_{k_1=k_2\ne k_3} 
	X_{k_1}^{\ell}[j] X_{k_2}^{\ell}[j + \Delta_1] X_{k_3}^{\ell}[j + \Delta_2]
}_{\text{(b)}}
\nonumber \\
&+\underbrace{ \sum_{j=0}^{L-1} \sum_{k_1=k_3\ne k_2} 
	X_{k_1}^{\ell}[j] X_{k_2}^{\ell}[j + \Delta_1] X_{k_3}^{\ell}[j + \Delta_2]
}_{\text{(c)}}
\nonumber \\
&+\underbrace{ \sum_{j=0}^{L-1} \sum_{k_2=k_3\ne k_1} 
	X_{k_1}^{\ell}[j] X_{k_2}^{\ell}[j + \Delta_1] X_{k_3}^{\ell}[j + \Delta_2]
}_{\text{(d)}}
\nonumber \\
&+\underbrace{ \sum_{j=0}^{L-1} \sum_{k_1=k_2=k_3} 
	X_{k_1}^{\ell}[j] X_{k_2}^{\ell}[j + \Delta_1] X_{k_3}^{\ell}[j + \Delta_2]
}_{\text{(e)}}.
%
\end{align}

For term (a), the expectation conditional on $M$ is:
%
\begin{align}
%
\sum_{j=0}^{L-1} M_\ell(M_\ell-1)(M_\ell-2)\M[j] \M[j+\Delta_1] \M[j+\Delta_2].
%
\end{align}
%
Using Lemma \ref{lem-choose}, the unconditional expectation of (a) is then:
%
\begin{align} \label{aaaa}
%
\gamma^3 \sum_{j=0}^{L-1} \M_1[j] \M_1[j+\Delta_1] \M_1[j+\Delta_2].
%
\end{align}

For term (b), the expectation conditional on $M$ is:
%
\begin{align}
%
\sum_{j=0}^{L-1} M_\ell (M_\ell - 1) \M_2[j,j+\Delta_1] \M_1[j + \Delta_2]
%
\end{align}
%
and then again using Lemma \ref{lem-choose} we get the expected value:
%
\begin{align} \label{bbbb}
%
\gamma^2 \sum_{j=0}^{L-1} \M_2[j,j+\Delta_1] \M_1[j+\Delta_2].
%
\end{align}

Similarly, the expected values of terms (c) and (d) are:
%
\begin{align} \label{cccc}
%
\gamma^2 \sum_{j=0}^{L-1} \M_2[j,j+\Delta_2] \M_1[j+\Delta_1].
%
\end{align}
%
and
%
\begin{align} \label{dddd}
%
\gamma^2 \sum_{j=0}^{L-1} \M_2[j+\Delta_1,j+\Delta_2] \M_1[j].
%
\end{align}

Finally, the expected value of term (e) is easily shown to be:
%
\begin{align} \label{eeee}
%
\gamma \sum_{j=0}^{L-1} \M_3[j,j+\Delta_1,j+\Delta_2].
%
\end{align}
%
This concludes the computation for Case \ref{case1}.

Moving onto Case \ref{case2}, we have $\ell_1 \equiv i_1 - j_1 = i_2 - j_2$, and also define $\ell_2 \equiv i_3 - j_3$. By definition, $\ell_1 \ne \ell_2$. The sum is:
%
\begin{align}
%
& \sum_{j_1=0}^{L-1} \sum_{j_3 \ne j_1 + \Delta_2}
\sum_{1 \le k_1,k_2 \le M_{\ell_1}} \sum_{k_3=1}^{M_{\ell_2}}
X_{k_1}^{\ell_1}[j_1] X_{k_2}^{\ell_1}[j_1 + \Delta_1] X_{k_3}^{\ell_2}[j_3]   
\nonumber \\
=& \sum_{j_1=0}^{L-1} \sum_{j_3 \ne j_1 + \Delta_2} \sum_{k_3=1}^{M_{\ell_2}}
\Bigg\{
\sum_{1 \le k_1 \ne k_2 \le M_{\ell_1}} 
X_{k_1}^{\ell_1}[j_1] X_{k_2}^{\ell_1}[j_1 + \Delta_1] X_{k_3}^{\ell_2}[j_3]
\nonumber \\
&       + \sum_{k_1=1}^{M_{\ell_1}} X_{k_1}^{\ell_1}[j_1] X_{k_1}^{\ell_1}[j_1 + \Delta_1] 
X_{k_3}^{\ell_2}[j_3]  \Bigg\}.
%
\end{align}
%
Taking expectations conditional on $M$, we then get:
%
\begin{align}
%
& \sum_{j_1=0}^{L-1} \sum_{j_3 \ne j_1 + \Delta_2} 
\Bigg(M_{\ell_1} (M_{\ell_1}-1) M_{\ell_2} \M_1[j_1] \M_1[j_1 + \Delta_1] \M_1[j_3]
\nonumber \\
&       + M_{\ell_1} M_{\ell_2} \M_2[j_1,j_1+\Delta_1] \M_1[j_3] \Bigg).
%
\end{align}

Taking expectations over $M$ and using Lemma \ref{lem-choose} then gives:
%
\begin{align}
%
& \gamma^3 \sum_{j_1=0}^{L-1} \sum_{j_3 \ne j_1 + \Delta_2} 
\M_1[j_1] \M_1[j_1 + \Delta_1] \M_1[j_3]
\label{ffff} \\
& + \gamma^2 \sum_{j_1=0}^{L-1} \sum_{j_3 \ne j_1 + \Delta_2}  
\M_2[j_1,j_1+\Delta_1] \M_1[j_3].
\label{gggg}
%
\end{align}

Similarly, Cases \ref{case3} and \ref{case4} give the expressions:
%
\begin{align}
%
& \gamma^3 \sum_{j_1=0}^{L-1} \sum_{j_2 \ne j_1 + \Delta_1} 
\M_1[j_1] \M_1[j_1 + \Delta_2] \M_1[j_2]
\label{hhhh} \\
& + \gamma^2 \sum_{j_1=0}^{L-1} \sum_{j_2 \ne j_1 + \Delta_1}  
\M_2[j_1,j_1+\Delta_2] \M_1[j_2]
\label{iiii}
%
\end{align}
%
and
%
\begin{align}
%
& \gamma^3 \sum_{j_2=0}^{L-1} \sum_{j_1 \ne j_2} 
\M_1[j_1] \M_1[j_2 + \Delta_1] \M_1[j_2 + \Delta_2]
\label{jjjj}\\
& + \gamma^2 \sum_{j_2=0}^{L-1} \sum_{j_1 \ne j_2}  
\M_2[j_2+\Delta_1,j_2+\Delta_2] \M_1[j_1].
\label{kkkk}
%
\end{align}

Finally, in Case \ref{case5} we have $i_1 - j_1$, $i_2 - j_2$, and $i_3 - j_3$ are all pairwise distinct. Consequently, the $X$ variables are always independent, and the expectation conditional on $M$ (letting $\ell_q = i_q - j_q$, $q=1,2,3$),
%
\begin{align}
%
\sum_{j_1,j_2,j_3} M_{\ell_1} M_{\ell_2} M_{\ell_3} \M_1[j_1] \M_1[j_2] \M_1[j_3];
%
\end{align}
%
since the $M_{\ell_q}$'s are pairwise independent, $q=1,2,3$, the expectation over $M$ then yields:
%
\begin{align} \label{llll}
%
\gamma^3 \sum_{j_1,j_2,j_3} \M_1[j_1] \M_1[j_2] \M_1[j_3].
%
\end{align}

Now we add all the terms from Cases \ref{case1} to \ref{case5}. Expressions \eqref{aaaa}, \eqref{ffff}, \eqref{hhhh}, \eqref{jjjj}, and \eqref{llll} sum to the expression:
%
\begin{align}
%
(\gamma \L_1)^3.
%
\end{align}

Note that this is obtained directly from the first moment. Expressions \eqref{bbbb}, \eqref{cccc}, \eqref{dddd}, \eqref{gggg},\eqref{iiii}, and \eqref{kkkk} sum to the expression:
%
\begin{align}
%
\gamma \L_1  \cdot 
( \gamma\L_2(\Delta_1) + \gamma\L_2(\Delta_2) + \gamma\L_2(\Delta_2-\Delta_1)).
%
\end{align}
%
Again, note that this is obtained directly from the first two moments. Finally, expression \eqref{eeee} is simply:
%
\begin{align}
%
\gamma \L_3(\Delta_1,\Delta_2)
%
\end{align}
%
which is the usual third-order auto-correlation.
